\documentclass[aspectratio=169]{beamer}
\input{../utils/beamerpreamble.tex}

\subtitle[Statistical Indicators]{Topic 02 - Point and Interval Indicators}

\begin{document}
\begin{CJK}{UTF8}{ipxm}

\begin{frame}
  \maketitle

  \vfill

  \hfill \tiny{Version 2021.2 (Updated \today)}
\end{frame}

\section{Intro}
\begin{frame}
  \begin{center}
    Part I - Intro
  \end{center}
\end{frame}

\begin{frame}{What is today's class about?}{Topics}
  \begin{itemize}
    \item Main topic: Indicators;
    \begin{itemize}
      \item A first look at data from experiments;
      \item Point Indicators: What our data tells us about the world?
      \item Interval Indicators: How sure are we of our data?
      \item Some statistical concepts;
    \end{itemize}\bigskip

    \item Some example code;\bigskip

    \item Feedback from the last class;
  \end{itemize}\bigskip

  Don't forget to bring your questions to the office hour!
\end{frame}

\input{03_indicators.tex}

\section{Outro}
\begin{frame}
  \begin{center}
    Part III - Outro
  \end{center}
\end{frame}

\begin{frame}{Summary}{Descriptive Statistics}

  Experiment Data can be used to \structure{estimate facts about the world}:
  \bigskip

  \begin{itemize}
    \item Point estimators: Sample Means, Variance, Correlation, etc.
    \begin{itemize}
      \item Give us specific information about the model we want to study
      \item "What is the average height of a student?"
      \item {\bf An estimator is not the real value!}
    \end{itemize}\bigskip

    \item Interval estimators: Confidence Interval, IQR, etc.
    \begin{itemize}
      \item Give us more information than point estimators.
      \item "How certain should I be about this point estimator".
      \item Size of interval estimator depends on the number of samples.
    \end{itemize}
  \end{itemize}
\end{frame}


\begin{frame}{Recommended Reading}
  \begin{itemize}
    \item \emph{D.W. Stockburger}, The Sampling Distribution. In: Introductory Statistics: Concepts, Models, and Applications -
    \url{http://psychstat3.missouristate.edu/Documents/IntroBook3/sbk17.htm}
    \item \emph{J.G. Ramírez}, Statistical Intervals: Confidence, Prediction, Enclosure: \url{https://git.io/v5ZFh}
    \item Crash Course Statistics Playlist, in particular videos \#3 to \#7: \url{https://www.youtube.com/playlist?list=PL8dPuuaLjXtNM_Y-bUAhblSAdWRnmBUcr}
  \end{itemize}
\end{frame}

\begin{frame}{Programming in R}
  The material for this week includes some coding examples. These examples are
  written in the {\bf R} language.\bigskip

  Although we will have an R tutorial in the future, you can read the
  following material to get yourself acquainted with R:\bigskip

  \begin{itemize}
    \item R for beginners: \url{https://cran.r-project.org/doc/contrib/Paradis-rdebuts_en.pdf}
    \item Rstudio: \url{https://rstudio.com}
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{../utils/lastslide.tex}
\end{CJK}
\end{document}
