---
title: "Data Analysis Example: Student Height"
subtitle: "Code Example -- Lecture 02"
author: "Claus Aranha"
---

# Introduction

This example analyses a survey from two classes of students about their height 
and weight. What can we learn about these students from this data?

# Data 

The data for this example was obtained from Felipe Campelo's "Analysis and 
Design of Experiments" course. and is available as a .csv file in the "rawdata"
subdirectory.

```{r}
students <- read.csv("rawdata/student_data.csv")
students # this is an R dataframe
```

> **Note:** When doing data analysis, it is always good to keep your raw 
(original) data files in a separate directory. Data analysis usually requires 
modifications to the data frame, so it is important to have the original, 
non-modified data somewhere handy.

# Simple Analysis and visualization:

What is the mean height and weight of these students?

```{r}
mean.height <- mean(students$Height.m)   # Student mean height
mean.weight <- mean(students$Weight.kg)   # Student mean weight

print(c(mean.height, mean.weight))
```

We are interested in the BMI (Body Mass Index) of the students. The BMI is 
sometimes used to indicate the relationship between the height and the 
weight of a student, and is given by the formula:

$BMI = \frac{Weight}{Height^2}$

Let's calculate the BMI and add it to our `students` dataframe.

```{r students_bmi}
  bmi = students$Weight.kg / (students$Height.m ** 2)
  students["BMI"] <- bmi
  students
```

The mean BMI of all the students is `r mean(students$BMI)` with an estimated 
error of `r sd(students$BMI)`.

# Figures and confidence interval:

When analyzing data, it is important to accompany point indicators (values) with
statistical intervals. Using **figures** also gives the reader a good intuition
of the relationship between all the values under study.

Let's understand the BMI of the students in this class a little bit more, by
plotting the 95% confidence interval for this estimator:

Let's plot the mean BMI of the students, as well as the 95% confidence interval for this estimator:

```{r students_plotbmi}
values <- students$BMI  # shorter name for variable
ci_alpha <- 0.05        # Alpha value for confidence interval. Confidence = 1 - alpha.
# calculating low and high bounds for ci
ci_low <- mean(values) + qt(ci_alpha/2, length(values)-1)*(sd(values)/sqrt(length(values))) 
ci_hi <- mean(values) + qt(1 - ci_alpha/2, length(values)-1)*(sd(values)/sqrt(length(values)))

# plot the values and the confidence interval of the mean
plot(students$BMI, ylim = c(0,40), xlab = "Student ID", ylab = "BMI")
title("BMI Values")
abline(h = mean(students$BMI))
abline(h = ci_low, lty=2)
abline(h = ci_hi, lty=2)
```

*TODO*: Print a prettier plot with ggplot.

# Investigating relationship between variables

Is there a difference in BMI between male and female students in this class?

```{r students_bmi_gender, echo=FALSE}
# ECHO=FALSE hides the code in the report

# Creating two separate data frames, one for males, one for females.
students_male <- students[which(students$Gender == "M"),]
students_female <- students[which(students$Gender == "F"),]

# Create a function to calculate confidence intervals.
ci_bounds <- function(x,alpha) {
  low <- mean(x) + qt(alpha/2, length(x)-1)*(sd(x)/sqrt(length(x))) 
  hi <- mean(x) + qt(1 - alpha/2, length(x)-1)*(sd(x)/sqrt(length(x)))
  return(c(low,hi))
}

# plot confidence intervals for male and female students
plot(students_male$BMI, ylim = c(0,40), xlim = c(0,40), xlab = "Student ID", ylab = "BMI", col="red")
par(new = TRUE)
plot(students_female$BMI, ylim = c(0,40), xlim = c(0,40), xlab = "Student ID", ylab = "BMI", col="blue")
title("BMI Values")

ci_male <- ci_bounds(students_male$BMI, 0.01)
abline(h = mean(students_male$BMI), col="red")
abline(h = ci_male, lty=2, col="red")
abline(h = ci_male[2], lty=2, col="red")

ci_female <- ci_bounds(students_female$BMI, 0.01)
abline(h = mean(students_female$BMI), col="blue")
abline(h = ci_female[1], lty=2, col="blue")
abline(h = ci_female[2], lty=2, col="blue")

# TODO: Add a legend to this plot, or make a pretties plot using ggplot.
```

The plot above shows that not only the mean BMI of the female students (blue line)
is lower, but the 95% confidence interval of the two measures has very little 
overlap. 

But what does the confidence interval means in this scenario? Since we have 
information about all the students in our class, the confidence interval does 
not tells us much, unless we are assuming that this class is a representative 
sample of a larger population. 

But at least now we know how to calculate a confidence interval!

# Other ways to observe data.

The histogram allows us to visualize some data by breaking it down into 
categories:

```{r students_bmi_histogram, echo=FALSE}
hist(students$BMI)
```

The boxplot gives us a good idea of the upper and lower limits in a sample,
as well as the position of the majorities of the values in the sample.

```{r students_bmi_boxplot, echo=FALSE}
boxplot(BMI ~ Course, data=students)  # In R, the "X~Y" notation means "variable X explained by variable Y"
title("Boxplot of BMI values of students depending on course")
```
