---
title: "Data Analysis Example: Coaxial Cables"
subtitle: "Code Example -- Lecture 02"
author: "Claus Aranha"
---

# Introduction

Let's assume a hypothetical factory that produces coaxial cables.
We are interested in measuring the resistance of the cables produced
by the factory.

First, let's create a hypothetical data set of cables produced by a 
process that follows a Gaussian distribution, with mean **50** and 
standard deviation **2**:

```{r coaxial_generate_data}
# Generating the data. 
set.seed(42)                    # Set PRNG seed (for reproducibility)
x<-rnorm(n=25, mean=50, sd=2)   # Draw 25 samples from N(mu=50,sigma=2)

x                               # Lists the resistances of each cable in the sample
```

So our hypothetical factory generated cables with the resistances above. 
That is our data for this example.

# Point estimators.

When we get data from an experiment, we are usually interested in calculating
*point estimators* that describe the characteristics of the process that 
generated that data.

So if we see a list of resistances for cables in a factory, we want to know
the average resistance of a "standard" cable from that factory. 

The **sample mean** is an unbiased estimator of the population mean, so we 
can use the **sample mean** from the data, to **estimate** the mean of the 
factory.

```{r coaxial_sample_mean}
sample_mean <- sum(x) / length(x) # Sample mean: sum of obs / sample size
sample_mean
```

Note that the **sample mean** (`r sample_mean`) is different from the 
**population mean** (`r 50`), because it is an **estimator**.

Since it is an estimator, the sample mean has an error. We can 
calculate the estimated error as follows:

```{r coaxial_estimated_error}
sample_mean_error <- sd(x)/sqrt(length(x)) # sample mean error: slide 25
sample_mean_error
```

If we collect a larget sample, our estimated error for the sample mean 
becomes smaller. This is one reason why we want to do many experiments:

```{r coaxial_sample_size}
csample_10 <- rnorm(n=10, mean=50, sd=2) # sample 10 cables
csample_25 <- rnorm(n=25, mean=50, sd=2) # sample 25 cables
csample_50 <- rnorm(n=50, mean=50, sd=2) # sample 50 cables

sme <- function(x) {
  sd(x) / sqrt(length(x))
}
```

For a sample with 10 cables, the sample mean is `r mean(csample_10)`, and the sample mean error is `r sme(csample_10)`.

For a sample with 25 cables, the sample mean is `r mean(csample_25)`, and the sample mean error is `r sme(csample_25)`.

For a sample with 50 cables, the sample mean is `r mean(csample_50)`, and the sample mean error is `r sme(csample_50)`.

